---
title: "Active-transport walking behavior"
output:
  pdf_document: default
  html_document: default
date: '2022-05-19'
---

# Active-transport walking and cycling behaviour in Canada


## Introduction 



## Data and Method

Data on walking and cycling trips to different destinations was derived from Canada’s General Social Survey (GSS) from 1986 to 2015. GSS is a social trends survey in order to monitor changes in the living conditions and well-being of Canadians and to provide information on specific social policy issues. one of the important information of GSS is about Time-use surveys. This survey collects information on all human activities and can therefore inform a broad range of policies. Statistics Canada has been conducting time-use surveys since 1986 at approximately five- to seven- year intervals, most recently in 2015.The GSS on time use employs a retrospective 24-hour time diary to collect information on respondents’ participation in, and time spent on, a wide variety of day-to-day activities. In addition, information is collected on the location where these activities occurred (e.g., at home, at work, etc.).in addition, This dataset contains travel time data of Many of the Census Metropolitan Areas (CMAs) and non-CMA areas all over Canada. CMAs are including St. John’s, Halifax, Saint John, Montreal, Quebec City, Toronto, Ottawa, Hamilton, Winnipeg, Regina, Saskatoon, Calgary, Edmonton, and Vancouver. and the non-CMA areas of each of the ten provinces were also grouped to form ten more strata.

walking and cycling data are categorized based on the three different destinations such as home, school and work, and others' homes. every episode includes, the start and end time of each event, duration,  weight, and concurrent contextual information, such as the location of the activities (e.g., home, work, schools,...). 


complete method


# Result



```{r load-packages, include=FALSE, cache=FALSE}
#load all packages:
library(fitdistrplus)
library(dplyr)
library(ggplot2)
library(readxl) 
library(splitstackshape)
library(tibble)
library(readr)
library(tabulate)
library(janitor)
library(kableExtra)
library(flextable)
```


```{r, include=FALSE, cache=FALSE}
#Read main file:
gss_e_2015 <- read.csv("D:/GSS Data/gss-e.csv")
gss_e_2005 <- read.csv("D:/First Article/Data/Time use_2005/gss-12M0019-E-2005-c-19-e_F1.csv")
gss_e_1998 <- read.csv("D:/First Article/Data/Time use_1998/gss-12M0012-E-1998-c-12e_F1.csv")
gss_e_1992 <- read.csv("D:/First Article/Data/Time use_1992/gss-12M0007-E-1992-c-7-ep_F1.csv")
gss_e_1986 <- read.csv("D:/First Article/Data/Time use_1986/gss-12M0002-E-1986-c-2-ep_F1.csv")
```


```{r creating dataset for walking 2015, include=FALSE, cache=FALSE}

# **walking 2015**
# Creating data of origins and destinations of walking trip 2015
inds = which(gss_e_2015$LOCATION == 315)
rows <- lapply(inds, function(x) (x-1):(x+1))
walking_2015 <- gss_e_2015 [unlist(rows),] %>% 
  dplyr::select(PUMFID:LOCATION) %>% 
  mutate(origin = lag(LOCATION, order_by = PUMFID)) %>% 
  mutate(destination = lead(LOCATION, order_by = PUMFID)) %>% 
  group_by(PUMFID) %>% 
  filter(LOCATION == 315) %>% 
  ungroup()

# change destination and origins to text column
walking_2015 <- walking_2015 %>%  filter(destination == 300 | destination == 301 | destination == 303 | destination == 302 | destination == 305 |destination == 306 | destination == 307 | destination == 309 | destination == 310 ) %>% filter(origin == 300 | origin == 301 | origin == 303 | origin == 302 | origin == 305 | origin == 306 | origin == 307 | origin == 309 | origin == 310) %>% 
  mutate(dest =
                     case_when(destination == 300 ~ "home", 
                               destination == 301 ~ "work or school",
                               destination == 303 ~ "other's home",
                               destination == 302 ~ "business",
                               destination == 305 ~ "outdoors",
                               destination == 306 ~ "Grocery store, other stores or mall",
                               destination == 306 ~ "Library, museum or theatre",
                               destination == 309 ~ "Restaurant, bar or club",
                               destination == 310 ~ "Place of worship")) %>% 
  mutate(orig =
                     case_when(origin == 300 ~ "home", 
                               origin == 301 ~ "work or school",
                               origin == 303 ~ "other's home",
                               origin == 302 ~ "business",
                               origin == 305 ~ "outdoors",
                               origin == 306 ~ "Grocery store, other stores or mall",
                               origin == 306 ~ "Library, museum or theatre",
                               origin == 309 ~ "Restaurant, bar or club",
                               origin == 310 ~ "Place of worship"))

walking_2015 <- walking_2015 %>%  dplyr::select(PUMFID, WGHT_EPI, DDAY, STARTIME:orig) 
walking_2015$YEAR <-  2015
walking_2015$MODE <-  "walking"
```

```{r}
summry (walking_2015$orig)
```




```{r creating dataset for cycling 2015, include=FALSE, cache=FALSE}

# cycling _2015
# Creating data of origins and destinations of cycling trip 2015
inds = which(gss_e_2015$LOCATION == 318)
rows <- lapply(inds, function(x) (x-1):(x+1))
cycling_2015 <- gss_e_2015[unlist(rows),] %>% 
  dplyr::select(PUMFID:LOCATION) %>% 
  mutate(origin = lag(LOCATION, order_by = PUMFID)) %>% 
  mutate(destination = lead(LOCATION, order_by = PUMFID)) %>% 
  group_by(PUMFID) %>% 
  filter(LOCATION == 318)

# change destination and origins to text column
cycling_2015 <- cycling_2015 %>%  filter(destination == 300 | destination == 301 | destination == 303 | destination == 302 | destination == 305 |destination == 306 | destination == 307 | destination == 309 | destination == 310 ) %>% filter(origin == 300 | origin == 301 | origin == 303 | origin == 302 | origin == 305 | origin == 306 | origin == 307 | origin == 309 | origin == 310) %>% 
  mutate(dest =
                     case_when(destination == 300 ~ "home", 
                               destination == 301 ~ "work or school",
                               destination == 303 ~ "other's home",
                               destination == 302 ~ "business",
                               destination == 305 ~ "outdoors",
                               destination == 306 ~ "Grocery store, other stores or mall",
                               destination == 306 ~ "Library, museum or theatre",
                               destination == 309 ~ "Restaurant, bar or club",
                               destination == 310 ~ "Place of worship")) %>% 
  mutate(orig =
                     case_when(origin == 300 ~ "home", 
                               origin == 301 ~ "work or school",
                               origin == 303 ~ "other's home",
                               origin == 302 ~ "business",
                               origin == 305 ~ "outdoors",
                               origin == 306 ~ "Grocery store, other stores or mall",
                               origin == 306 ~ "Library, museum or theatre",
                               origin == 309 ~ "Restaurant, bar or club",
                               origin == 310 ~ "Place of worship"))

cycling_2015 <- cycling_2015 %>%  dplyr::select(PUMFID, WGHT_EPI, DDAY, STARTIME:orig) 
cycling_2015$YEAR <-  2015
cycling_2015$MODE <-  "cycling"
```

```{r creating dataset for walking 2005, include=FALSE, cache=FALSE}

# **walking 2005**
# Creating data of origins and destinations of walking trip 2005
inds = which(gss_e_2005$PLACE == 14)
rows <- lapply(inds, function(x) (x-1):(x+1))
walking_2005 <- gss_e_2005 [unlist(rows),] %>% 
  dplyr::select(RECID:PLACE) %>% 
  mutate(origin = lag(PLACE, order_by = RECID)) %>% 
  mutate(destination = lead(PLACE, order_by = RECID)) %>% 
  group_by(RECID) %>% 
  filter(PLACE == 14) %>% 
  ungroup()

# change destination and origins to text column
walking_2005 <- walking_2005 %>%  filter(destination == 1 | destination == 2 | destination == 3 | destination == 8 | destination == 4 | destination == 5 | destination == 6 | destination == 7 | destination == 9 | destination == 10 ) %>% filter(origin == 1 | origin == 2 | origin == 3 | origin == 8 | origin == 4 | origin == 5 | origin == 6 | origin == 7 | origin == 9 | origin == 10) %>% 
  mutate(dest =
                     case_when(destination == 1 ~ "home", 
                               destination == 2 ~ "work or school",
                               destination == 3 ~ "other's home",
                               destination == 8 ~ "work or school",
                               destination == 4 ~ "Restaurant, bar or club",
                               destination == 5 ~ "Place of worship",
                               destination == 6 ~ "Grocery store, other stores or mall",
                               destination == 7 ~ "Grocery store, other stores or mall",
                               destination == 9 ~ "outdoors",
                               destination == 10 ~ "Library, museum or theatre")) %>% 
  mutate(orig =
                     case_when(origin == 1 ~ "home", 
                               origin == 2 ~ "work or school",
                               origin == 3 ~ "other's home",
                               origin == 8 ~ "work or school",
                               origin == 4 ~ "Restaurant, bar or club",
                               origin == 5 ~ "Place of worship",
                               origin == 6 ~ "Grocery store, other stores or mall",
                               origin == 7 ~ "Grocery store, other stores or mall",
                               origin == 9 ~ "outdoors",
                               origin == 10 ~ "Library, museum or theatre"))

walking_2005 <- walking_2005 %>% dplyr::rename(
    PUMFID = RECID,
    STARTMIN = STARMIN,
    LOCATION = PLACE)

walking_2005 <- walking_2005 %>%  dplyr::select(PUMFID, WGHT_EPI, DDAY, STARTIME:orig) 
walking_2005$YEAR <-  2005
walking_2005$MODE <-  "walking"
```

```{r creating dataset for cycling 2005, include=FALSE, cache=FALSE}

# **CYCLING 2005**
# Creating data of origins and destinations of cycling trip 2005
inds = which(gss_e_2005$PLACE == 17)
rows <- lapply(inds, function(x) (x-1):(x+1))
cycling_2005 <- gss_e_2005 [unlist(rows),] %>% 
  dplyr::select(RECID:PLACE) %>% 
  mutate(origin = lag(PLACE, order_by = RECID)) %>% 
  mutate(destination = lead(PLACE, order_by = RECID)) %>% 
  group_by(RECID) %>% 
  filter(PLACE == 17) %>% 
  ungroup()

# change destination and origins to text column
cycling_2005 <- cycling_2005 %>%  filter(destination == 1 | destination == 2 | destination == 3 | destination == 8 ) %>% filter(origin == 1 | origin == 2 | origin == 3 | origin == 8 ) %>% 
  mutate(dest =
                     case_when(destination == 1 ~ "home", 
                               destination == 2 ~ "work or school",
                               destination == 3 ~ "other's home",
                               destination == 8 ~ "work or school")) %>% 
  mutate(orig =
                     case_when(origin == 1 ~ "home", 
                               origin == 2 ~ "work or school",
                               origin == 3 ~ "other's home",
                               origin == 8 ~ "work or school"))

cycling_2005 <- cycling_2005 %>% dplyr::rename(
    PUMFID = RECID,
    STARTMIN = STARMIN,
    LOCATION = PLACE) %>%  dplyr::select(PUMFID, WGHT_EPI, DDAY, STARTIME:orig) 
cycling_2005$YEAR <-  2005
cycling_2005$MODE <-  "cycling"
```

```{creating dataset for walking 1998, include=FALSE, cache=FALSE}

# **walking 1998**
# Creating data of origins and destinations of walking trip 1998
inds = which(gss_e_1998$PLACE == 7)
rows <- lapply(inds, function(x) (x-1):(x+1))
walking_1998 <- gss_e_1998 [unlist(rows),] %>% 
  dplyr::select(RECID:PLACE) %>% 
  mutate(origin = lag(PLACE, order_by = RECID)) %>% 
  mutate(destination = lead(PLACE, order_by = RECID)) %>% 
  group_by(RECID) %>% 
  filter(PLACE == 7) %>% 
  ungroup()

# change destination and origins to text column
walking_1998 <- walking_1998 %>%  filter(destination == 1 | destination == 2 | destination == 3) %>% filter(origin == 1 | origin == 2 | origin == 3) %>% 
  mutate(dest =
                     case_when(destination == 1 ~ "home", 
                               destination == 2 ~ "work or school",
                               destination == 3 ~ "other's home")) %>% 
  mutate(orig =
                     case_when(origin == 1 ~ "home", 
                               origin == 2 ~ "work or school",
                               origin == 3 ~ "other's home"))

walking_1998 <- walking_1998 %>% dplyr::rename(
    PUMFID = RECID,
    STARTMIN = STARMIN,
    WGHT_EPI = WGHTEPI,
    LOCATION = PLACE) %>%  dplyr::select(PUMFID, WGHT_EPI, DDAY, STARTIME:orig) 
walking_1998$YEAR <-  1998
walking_1998$MODE <-  "walking"
```

```{r creating dataset for cycling 1998, include=FALSE, cache=FALSE}

# **Cycling 1998**
# Creating data of origins and destinations of cycling trip 1998
inds = which(gss_e_1998$PLACE == 9)
rows <- lapply(inds, function(x) (x-1):(x+1))
cycling_1998 <- gss_e_1998 [unlist(rows),] %>% 
  dplyr::select(RECID:PLACE) %>% 
  mutate(origin = lag(PLACE, order_by = RECID)) %>% 
  mutate(destination = lead(PLACE, order_by = RECID)) %>% 
  group_by(RECID) %>% 
  filter(PLACE == 9) %>% 
  ungroup()

# change destination and origins to text column
cycling_1998 <- cycling_1998 %>%  filter(destination == 1 | destination == 2 | destination == 3) %>% filter(origin == 1 | origin == 2 | origin == 3) %>% 
  mutate(dest =
                     case_when(destination == 1 ~ "home", 
                               destination == 2 ~ "work or school",
                               destination == 3 ~ "other's home")) %>% 
  mutate(orig =
                     case_when(origin == 1 ~ "home", 
                               origin == 2 ~ "work or school",
                               origin == 3 ~ "other's home"))

cycling_1998 <- cycling_1998 %>% dplyr::rename(
    PUMFID = RECID,
    STARTMIN = STARMIN,
    WGHT_EPI = WGHTEPI,
    LOCATION = PLACE) %>%  dplyr::select(PUMFID, WGHT_EPI, DDAY, STARTIME:orig) 
cycling_1998$YEAR <-  1998
cycling_1998$MODE <-  "cycling"
```

```{r creating dataset for walking 1992, include=FALSE, cache=FALSE}
# **walking 1992**
# Creating data of origins and destinations of walking trip 1992
inds = which(gss_e_1992$PLACE == 7)
rows <- lapply(inds, function(x) (x-1):(x+1))
walking_1992 <- gss_e_1992 [unlist(rows),] %>% 
  dplyr::select(SEQNUM, DDAY, NOEPISO, ACTCODE, STARTIME, ENDTIME, DURATION, PLACE, TIMEWGT) %>% 
  mutate(origin = lag(PLACE, order_by = SEQNUM)) %>% 
  mutate(destination = lead(PLACE, order_by = SEQNUM)) %>% 
  group_by(SEQNUM) %>% 
  filter(PLACE == 7) %>% 
  ungroup()

# change NA's Weight to 1 


# change destination and origins to text column
walking_1992 <- walking_1992 %>%  filter(destination == 1 | destination == 2 | destination == 3) %>% filter(origin == 1 | origin == 2 | origin == 3) %>% 
  mutate(dest =
                     case_when(destination == 1 ~ "home", 
                               destination == 2 ~ "work or school",
                               destination == 3 ~ "other's home")) %>% 
  mutate(orig =
                     case_when(origin == 1 ~ "home", 
                               origin == 2 ~ "work or school",
                               origin == 3 ~ "other's home"))

walking_1992 <- walking_1992 %>% dplyr::rename(
    PUMFID = SEQNUM,
    LOCATION = PLACE, WGHT_EPI = TIMEWGT) %>%  dplyr::select(PUMFID, DDAY, STARTIME:orig) 
walking_1992$YEAR <-  1992
walking_1992$MODE <-  "walking"
walking_1992$STARTMIN <-  1
walking_1992$ENDMIN <-  1
```

```{r creating dataset for cycling 1992, include=FALSE, cache=FALSE}

# **Cycling 1992**
# Creating data of origins and destinations of cycling trip 1992
inds = which(gss_e_1992$PLACE == 9)
rows <- lapply(inds, function(x) (x-1):(x+1))
cycling_1992 <- gss_e_1992 [unlist(rows),] %>% 
  dplyr::select(SEQNUM, DDAY, NOEPISO, ACTCODE, STARTIME, ENDTIME, DURATION, PLACE, TIMEWGT) %>% 
  mutate(origin = lag(PLACE, order_by = SEQNUM)) %>% 
  mutate(destination = lead(PLACE, order_by = SEQNUM)) %>% 
  group_by(SEQNUM) %>% 
  filter(PLACE == 9) %>% 
  ungroup()

# change destination and origins to text column
cycling_1992 <- cycling_1992 %>%  filter(destination == 1 | destination == 2 | destination == 3) %>% filter(origin == 1 | origin == 2 | origin == 3) %>% 
  mutate(dest =
                     case_when(destination == 1 ~ "home", 
                               destination == 2 ~ "work or school",
                               destination == 3 ~ "other's home")) %>% 
  mutate(orig =
                     case_when(origin == 1 ~ "home", 
                               origin == 2 ~ "work or school",
                               origin == 3 ~ "other's home"))

cycling_1992 <- cycling_1992 %>% dplyr::rename(
    PUMFID = SEQNUM,
    LOCATION = PLACE,WGHT_EPI = TIMEWGT) %>%  dplyr::select(PUMFID, DDAY, STARTIME:orig) 
cycling_1992$YEAR <-  1992
cycling_1992$MODE <-  "cycling"
cycling_1992$STARTMIN <-  1
cycling_1992$ENDMIN <-  1

```

```{r creating dataset for walking 1986, include=FALSE, cache=FALSE}

# **walking 1986**
# Creating data of origins and destinations of walking trip 1986
inds = which(gss_e_1986$PLACE == 7)
rows <- lapply(inds, function(x) (x-1):(x+1))
walking_1986 <- gss_e_1986 [unlist(rows),] %>% 
  dplyr::select(SEQNUM, DDAY, NO_EPISO, ACT_CODE, STRTTIME, ENDTIME, DURATION, PLACE, FWGT_MS) %>% 
  mutate(origin = lag(PLACE, order_by = SEQNUM)) %>% 
  mutate(destination = lead(PLACE, order_by = SEQNUM)) %>% 
  group_by(SEQNUM) %>% 
  filter(PLACE == 7) %>% 
  ungroup()
summary(walking_1986$PLACE)

# change destination and origins to text column
walking_1986 <- walking_1986 %>%  filter(destination == 1 | destination == 2 | destination == 3) %>% filter(origin == 1 | origin == 2 | origin == 3) %>% 
  mutate(dest =
                     case_when(destination == 1 ~ "home", 
                               destination == 2 ~ "work or school",
                               destination == 3 ~ "other's home")) %>% 
  mutate(orig =
                     case_when(origin == 1 ~ "home", 
                               origin == 2 ~ "work or school",
                               origin == 3 ~ "other's home"))

walking_1986 <- walking_1986 %>% dplyr::rename(
    PUMFID = SEQNUM,
    LOCATION = PLACE, WGHT_EPI = FWGT_MS, STARTIME = STRTTIME) %>%  dplyr::select(PUMFID, DDAY, STARTIME:orig) 
walking_1986$YEAR <-  1986
walking_1986$MODE <-  "walking"
walking_1986$STARTMIN <-  1
walking_1986$ENDMIN <-  1
```



```{r, include=FALSE,  cache=FALSE}
# creating an integrated data frame from 1992 to 2015
trip <- rbind(walking_2015, cycling_2015, walking_2005, cycling_2005, walking_1998, cycling_1998, walking_1992, cycling_1992, walking_1986)
```


4.1 Walking and cycling duration 

Table 1. shows home is the most destination for walking trips.

```{r}
Table_1 <- trip %>% filter(MODE == "walking") %>% 
  group_by(dest, YEAR) %>%
 summarise(across(.cols = c(DURATION),  # columns
                   .fns =  list("mean" = mean, "Maximum" = max, "Median" = median),                               # 
                   na.rm=T))  
  walking_2015 %>%  count(dest) %>%
  mutate(percent = scales::percent(n / sum(n)))
  
   walking_2005 %>%  count(dest) %>%
  mutate(percent = scales::percent(n / sum(n)))
   
    walking_1998 %>%  count(dest) %>%
  mutate(percent = scales::percent(n / sum(n)))
     
    walking_1992 %>%  count(dest) %>%
  mutate(percent = scales::percent(n / sum(n)))
  
  kable(Table_1) %>%
  kable_styling(latex_options = c("scale_down")) 
#kbl(caption = "Summary statistics on walking trips in 2015") %>% kable_classic(full_width = F, html_font = "Cambria") %>%  add_header_above(c(" " = 1, "Duration" = 3)) 
```
 



Table 2. shows home is the most destination for cycling trips.

```{r}
Table_1 <- trip %>% 
  group_by(dest, YEAR) %>%
 summarise(across(.cols = c(DURATION),  # columns
                   .fns =  list("mean" = mean, "Maximum" = max, "Median" = median),                               # 
                   na.rm=T))  
  cycling_2015 %>%  count(dest) %>%
  mutate(percent = scales::percent(n / sum(n)))
  cycling_2005 %>%  count(dest) %>%
  mutate(percent = scales::percent(n / sum(n)))
  cycling_1998 %>%  count(dest) %>%
  mutate(percent = scales::percent(n / sum(n)))
  cycling_1992 %>%  count(dest) %>%
  mutate(percent = scales::percent(n / sum(n)))
  
  kable(Table_1) %>%
  kable_styling(latex_options = c("scale_down")) 
#  kbl(caption = "Summary statistics on walking trips in 2015") %>% kable_classic(full_width = F, html_font = "Cambria") %>%  add_header_above(c(" " = 1, "Duration" = 3)) 
```




percentage of walking trip during 1986- 2015 :

```{r}
crosstab <- xtabs(~ orig + dest, data = walking_2015)
round(100 * prop.table(crosstab, 1), 2)

crosstab <- xtabs(~ orig + dest, data = walking_2005)
round(100 * prop.table(crosstab, 1), 2)

crosstab <- xtabs(~ orig + dest, data = walking_1998)
round(100 * prop.table(crosstab, 1), 2)

crosstab <- xtabs(~ orig + dest, data = walking_1992)
round(100 * prop.table(crosstab, 1), 2)

crosstab <- xtabs(~ orig + dest, data = walking_1986)
round(100 * prop.table(crosstab, 1), 2)
```

percentage of cycling trip during 1992- 2015 :

```{r}
crosstab <- xtabs(~ orig + dest, data = cycling_2015)
round(100 * prop.table(crosstab, 1), 2)

crosstab <- xtabs(~ orig + dest, data = cycling_2005)
round(100 * prop.table(crosstab, 1), 2)

crosstab <- xtabs(~ orig + dest, data = cycling_1998)
round(100 * prop.table(crosstab, 1), 2)

crosstab <- xtabs(~ orig + dest, data = cycling_1992)
round(100 * prop.table(crosstab, 1), 2)

```



##calculating impedance function for walking trip during 1992- 2015:

```{r}
# plotting histograms
trip_w <- trip %>%  filter(MODE == "walking")
ggplot(trip_w, aes(x = DURATION)) + geom_histogram() + facet_grid(.~YEAR)


# plotting weighted histograms
trip_w <- trip %>%  filter(MODE == "walking")
ggplot(trip_w, aes(x = DURATION, weight = WGHT_EPI)) + geom_histogram() 

#hist(trip$DURATION %>% unlist() %>% as.numeric(), breaks=100) 
# a skew vs. kurtois graph
descdist(walking_2015$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
descdist(walking_2005$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
descdist(walking_1998$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
descdist(walking_1992$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
```


```{r}
#Let's test out different models for our walk trips, we see that lnorm or gamma will likely be the best fit according to the graph above. using fitdist function to fit a distribution using the default maximum likelihood estimation method and Nelder-Mead method for direct optimization

exp_w_2015_ <- fitdistrplus::fitdist(data=walking_2015$DURATION %>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(walking_2015$WGHT_EPI))

exp_w_2005_ <- fitdistrplus::fitdist(data=walking_2005$DURATION %>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(walking_2005$WGHT_EPI))

exp_w_1998_ <- fitdistrplus::fitdist(data=walking_1998$DURATION %>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(walking_1998$WGHT_EPI))

exp_w_1992_ <- fitdistrplus::fitdist(data=walking_1992$DURATION %>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(walking_1992$WGHT_EPI))

exp_w_1986_ <- fitdistrplus::fitdist(data=walking_1986$DURATION %>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(walking_1986$WGHT_EPI))

```

```{r}
summary(walking_1992$WGHT_EPI)
```


 we will pick EXPONENTIAL function for walk trips! 

```{r}
walking_2015_exp <- exp_w_2015_
walking_2005_exp <- exp_w_2005_
walking_1998_exp <- exp_w_1998_
walking_1992_exp <- exp_w_1992_
walking_1986_exp <- exp_w_1986_
summary(exp_w_2015_, exp_w_2005_, exp_w_1998_, exp_w_1992_, exp_w_1986_ )
```

now populate our dataframe with the impedance value (i.e. travel cost) based on their travel time. and then creating a new column based on multiple weight in value of impedance function. 

```{r}
walking_2015 <- walking_2015 %>%
 mutate(f = dexp(DURATION, walking_2015_exp$estimate["rate"])) 
summary(walking_2015$f)

walking_2005 <- walking_2005 %>%
 mutate(f = dexp(DURATION, walking_2005_exp$estimate["rate"])) 
summary(walking_2005$f)

walking_1998 <- walking_1998 %>%
 mutate(f = dexp(DURATION, walking_1998_exp$estimate["rate"])) 
summary(walking_1998$f)

walking_1992 <- walking_1992 %>%
mutate(f = dexp(DURATION, walking_1992_exp$estimate["rate"])) 
summary(walking_1992$f)

walking_1992 <- walking_1986 %>%
mutate(f = dexp(DURATION, walking_1986_exp$estimate["rate"])) 
summary(walking_1986$f)

# creating an integrated data frame from 1992 to 2015
trip_w_f <- rbind(walking_2015, walking_2005, walking_1998, walking_1992 , walking_1986)
```

** Impedance function for different destination **

```{r}
ggplot() + geom_line(data = trip_w_f, aes(x = DURATION, y = f), linetype="dashed") + facet_grid(dest~YEAR) 
```


##calculating impedance function for cycling trip during 1992- 2015:

repeat raw based on the weight column:
```{r}
#rcycling_2015 <- data.frame(lapply(cycling_2015, rep, cycling_2015$WGHT_EPI))
#rcycling_2005 <- data.frame(lapply(cycling_2005, rep, cycling_2005$WGHT_EPI))
#rcycling_1998 <- data.frame(lapply(cycling_1998, rep, cycling_1998$WGHT_EPI))
#rcycling_1992 <- data.frame(lapply(cycling_1992, rep, cycling_1992$WGHT_EPI))
trip_c <-  rbind(cycling_2015, cycling_2005, cycling_1998, cycling_1992)
```


```{r}
# plotting histograms
ggplot(trip_c, aes(x = DURATION)) + geom_histogram() + facet_grid(.~YEAR)

# plotting weighted histograms
ggplot(trip_c, aes(x = DURATION, weight = WGHT_EPI)) + geom_histogram() + facet_grid(.~YEAR)

#hist(trip$DURATION %>% unlist() %>% as.numeric(), breaks=100) 
# a skew vs. kurtois graph
descdist(cycling_2015$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
descdist(cycling_2005$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
descdist(cycling_1998$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
descdist(cycling_1992$DURATION %>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
```

```{r}
#Let's test out different models for our walk trips, we see that lnorm or gamma will likely be the best fit according to the graph above. using fitdist function to fit a distribution using the default maximum likelihood estimation method and Nelder-Mead method for direct optimization

exp_c_2015_ <- fitdistrplus::fitdist(data=cycling_2015$DURATION %>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(cycling_2015$WGHT_EPI))
exp_c_2005_ <- fitdistrplus::fitdist(data=cycling_2005$DURATION %>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(cycling_2005$WGHT_EPI))
exp_c_1998_ <- fitdistrplus::fitdist(data=cycling_1998$DURATION %>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(cycling_1998$WGHT_EPI))
exp_c_1992_ <- fitdistrplus::fitdist(data=cycling_1992$DURATION %>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead", weights = round(cycling_1992$WGHT_EPI))


```

we will pick EXPONENTIAL function for cycling trips! 

```{r}
cycling_2015_exp <- exp_c_2015_
cycling_2005_exp <- exp_c_2005_
cycling_1998_exp <- exp_c_1998_
cycling_1992_exp <- exp_c_1992_
summary(exp_c_2015_, exp_c_2005_, exp_c_1998_, exp_c_1992_ )
```

now populate our dataframe with the impedance value (i.e. travel cost) based on their travel time. and then creating a new column based on multiple weight in value of impedance function. 

```{r}
cycling_2015 <- cycling_2015 %>%
 mutate(f = dexp(DURATION, cycling_2015_exp$estimate["rate"])) 
summary(cycling_2015$f)

cycling_2005 <- cycling_2005 %>%
 mutate(f = dexp(DURATION, cycling_2005_exp$estimate["rate"])) 
summary(cycling_2005$f)

cycling_1998 <- cycling_1998 %>%
 mutate(f = dexp(DURATION, cycling_1998_exp$estimate["rate"])) 
summary(cycling_1998$f)

cycling_1992 <- cycling_1992 %>%
mutate(f = dexp(DURATION, cycling_1992_exp$estimate["rate"])) 
summary(cycling_1992$f)

# creating an integrated data frame from 1992 to 2015
trip_c_f <- rbind(cycling_2015, cycling_2005, cycling_1998, cycling_1992)
```

** Impedance function for different destination **

```{r}
trip_c_f <- trip_c_f %>%  filter(MODE == "cycling")
ggplot() + geom_line(data = trip_c_f, aes(x = DURATION, y = f), linetype="dashed") + facet_grid(dest~YEAR) 
```

```{r}
# change destination and origins to text column
inds = which(gss_e_2015$LOCATION == 315)
rows <- lapply(inds, function(x) (x-1):(x+1))
walking_2015_1 <- gss_e_2015 [unlist(rows),] %>% 
  dplyr::select(PUMFID:LOCATION) %>% 
  mutate(origin = lag(LOCATION, order_by = PUMFID)) %>% 
  mutate(destination = lead(LOCATION, order_by = PUMFID)) %>% 
  group_by(PUMFID) %>% 
  filter(LOCATION == 315) %>% 
  ungroup()


walking_2015_1 <- walking_2015_1 %>% 
  mutate(dest = case_when(destination == 300 ~ "home", 
                               destination == 301 ~ "work or school",
                               destination == 303 ~ "other's home",
                               destination == 302 ~ "business",
                               destination == 304 ~ "neighbourhood",
                               destination == 305 ~ "Outdoors",
                               destination == 306 ~ "Grocery",
                               destination == 307 ~ "Library",
                               destination == 308 ~ "Sports",
                               destination == 309 ~ "Restaurant")) %>% 
  mutate(orig =
                     case_when(origin == 300 ~ "home", 
                               origin == 301 ~ "work or school",
                               origin == 303 ~ "other's home",
                               origin == 302 ~ "business",
                               origin == 304 ~ "neighbourhood",
                               origin == 305 ~ "Outdoors",
                               origin == 306 ~ "Grocery",
                               origin == 307 ~ "Library",
                               origin == 308 ~ "Sports",
                              origin == 309 ~ "Restaurant"))
  
                               
walking_2015_1 %>%  count(dest) %>%
mutate(percent = scales::percent(n / sum(n)))                           

```



```{r}
Table_1 <- walking_2015_1  %>% 
 summarise(across(.cols = c(DURATION),  # columns
                   .fns =  list("mean" = mean, "Maximum" = max, "Median" = median),                               # 
                   na.rm=T)) 
Table_2 <- walking_2005  %>% 
 summarise(across(.cols = c(DURATION),  # columns
                   .fns =  list("mean" = mean, "Maximum" = max, "Median" = median),                               # 
                   na.rm=T))
Table_3 <- walking_1998  %>% 
 summarise(across(.cols = c(DURATION),  # columns
                   .fns =  list("mean" = mean, "Maximum" = max, "Median" = median),                               # 
                   na.rm=T))
Table_4 <- walking_1992  %>% 
 summarise(across(.cols = c(DURATION),  # columns
                   .fns =  list("mean" = mean, "Maximum" = max, "Median" = median),                               # 
                   na.rm=T))
Table_5 <- walking_1986  %>% 
 summarise(across(.cols = c(DURATION),  # columns
                   .fns =  list("mean" = mean, "Maximum" = max, "Median" = median),                               # 
                   na.rm=T))
kable(Table_1) %>%
  kable_styling(latex_options = c("scale_down")) 
kable(Table_2) %>%
  kable_styling(latex_options = c("scale_down")) 
kable(Table_3) %>%
  kable_styling(latex_options = c("scale_down")) 
kable(Table_4) %>%
  kable_styling(latex_options = c("scale_down")) 
kable(Table_5) %>%
  kable_styling(latex_options = c("scale_down"))



```




```{r}
Table_1 <- cycling_2015_1  %>% 
 summarise(across(.cols = c(DURATION),  # columns
                   .fns =  list("mean" = mean, "Maximum" = max, "Median" = median),                               # 
                   na.rm=T)) 
Table_2 <- cycling_2005  %>% 
 summarise(across(.cols = c(DURATION),  # columns
                   .fns =  list("mean" = mean, "Maximum" = max, "Median" = median),                               # 
                   na.rm=T))
Table_3 <- cycling_1998  %>% 
 summarise(across(.cols = c(DURATION),  # columns
                   .fns =  list("mean" = mean, "Maximum" = max, "Median" = median),                               # 
                   na.rm=T))
Table_4 <- cycling_1992  %>% 
 summarise(across(.cols = c(DURATION),  # columns
                   .fns =  list("mean" = mean, "Maximum" = max, "Median" = median),                               # 
                   na.rm=T))

kable(Table_1) %>%
  kable_styling(latex_options = c("scale_down")) 
kable(Table_2) %>%
  kable_styling(latex_options = c("scale_down")) 
kable(Table_3) %>%
  kable_styling(latex_options = c("scale_down")) 
kable(Table_4) %>%
  kable_styling(latex_options = c("scale_down")) 




```
