---
title: "R Notebook"
output: html_notebook
---

Begin by loading the packages used in this notebook: 
```{r}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(here) # A Simpler Way to Find Your Files
library(readxl) # Read Excel Files
library(tidyr) # Tidy Messy Data
library(usethis) # Automate Package and Project Setup
```


## Source Files

In Canada, there is a survey program run by Statistics Canada called General Social Surveys or GSS for short. GSS collects data on social trends in order to track changes in Canadians' living conditions and well-being over time, as well as to provide immediate information on specific social policy issues of current or emerging interest. This survey tracks changes in time use to better understand how Canadians spend and manage their time, as well as what factors contribute to their happiness and stress .This program in Canada was established in 1985 as a series of independent, annual, cross-sectional surveys, each covering one topic in depth. Caregiving, families, time use, social identity, volunteering, and victimization are among the current GSS themes.

Each of the six survey themes listed above is thoroughly repeated every 5 years. In addition to the core topic, each cycle includes new content that addresses emerging, policy-relevant issues. In addition, each survey collects detailed socio-demographic information such as age, gender, education, religion, ethnicity, income, and so on.

Time-use surveys gather data on all human activities and can thus inform a wide range of policies. In particular, three key themes have been identified as essential for informed policymaking and for which no other data sources are adequate: unpaid work and non-market production; well-being; and gender equality. Leisure time, work-life balance, health, commuting, culture, and sports are among the other topics covered by time use surveys. In addition, statistics Canada has been conducting time-use surveys at five- to seven-year intervals since 1986, most recently in 2015 (1986,1992,1998, 2005 ,2010 and 2015). This means that there is information about time use patterns covering a period of so many years.
The GSS on time use collects information on respondents' participation in, and time spent on, a wide range of day-to-day activities using a retrospective 24-hour time diary. Furthermore, information is gathered on the location of these activities (e.g., at home, at work, etc.) and, for non-personal activities, the people who were present with the respondent at the time of the activity.


The surveys (obtained from [here](http://odesi2.scholarsportal.info/webview/)) consist of two files: a main file and an episode file. These are their characteristics.

#The main file...contains info about respondents and their socio-economic...blah blah

#The episode file...contains blah blah

For the purpose of this notebook, the following definitions are important:

| **Code** 	| **Location**                             	| **Year** 	| **Description**                                                 	|
|----------	|------------------------------------------	|----------	|-----------------------------------------------------------------	|
|    300   	| At home or on property                   	|   2015   	| Walking (code = 315) and Cycling (code = 318) to this location  	|
|    301   	| At place of work or school               	|   2015   	| Walking (code = 315) and Cycling (code = 318) to this location  	|
|   302    	| Away on business                         	|   2015   	| Walking (code = 315) and Cycling (code = 318) to this location  	|
|    303   	| At someone else's home or property       	|   2015   	| Walking (code = 315) and Cycling (code = 318) to this location  	|
|    304   	| In the neighbourhood                     	|   2015   	| Walking (code = 315) and Cycling (code = 318) to this location  	|
|    305   	| Outdoors                                 	|   2015   	| Walking (code = 315) and Cycling (code = 318) to this location  	|
|    306   	| Grocery store, other stores or mall      	|   2015   	| Walking (code = 315) and Cycling (code = 318) to this location  	|
|    307   	| Library, museum or theatre               	|   2015   	| Walking (code = 315) and Cycling (code = 318) to this location  	|
|    308   	| Sports centre, field or arena            	|   2015   	| Walking (code = 315) and Cycling (code = 318) to this location  	|
|    309   	| Restaurant, bar or club                  	|   2015   	| Walking (code = 315) and Cycling (code = 318) to this location  	|
|    310   	| Place of worship                         	|   2015   	| Walking (code = 315) and Cycling (code = 318) to this location  	|
|    311   	| Medical, dental or other health   clinic 	|   2015   	| Walking (code = 315) and Cycling (code = 318) to this location  	|
|    312   	| Elsewhere                                	|   2015   	| Walking (code = 315) and Cycling (code = 318) to this location  	|
|     1    	| R's home                                 	|   2010   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     2    	| Work place                               	|   2010   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     3    	| Someone else's home                      	|   2010   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     4    	| Restaurant/bar                           	|   2010   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     5    	| Place of worship                         	|   2010   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     6    	| Grocery store                            	|   2010   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     7    	| Other store/Mall                         	|   2010   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     8    	| School                                   	|   2010   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     9    	| Outdoors away from   home                	|   2010   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|    10    	| Library                                  	|   2010   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|    11    	| Other place                              	|   2010   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     1    	| Respondent's home                        	|   2005   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     2    	| Work place                               	|   2005   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     3    	| Someone else's home                      	|   2005   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     4    	| Restaurant/bar                           	|   2005   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     5    	| Place of worship                         	|   2005   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     6    	| Grocery store                            	|   2005   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     7    	| Other store/Mall                         	|   2005   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     8    	| School                                   	|   2005   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|     9    	| Outdoors away from home                  	|   2005   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|    10    	| Library                                  	|   2005   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|    11    	| Other place                              	|   2005   	| Walking (code = 14) and Cycling (code = 17) to this location    	|
|    1     	| Respondent's home                        	|   1998   	| Walking (code = 7) and Cycling (code = 9) to this location      	|
|     2    	| Work place                               	|   1998   	| Walking (code = 7) and Cycling (code = 9) to this location      	|
|     3    	| Someone else's home                      	|   1998   	| Walking (code = 7) and Cycling (code = 9) to this location      	|
|     4    	| Other plourhood                          	|   1998   	| Walking (code = 7) and Cycling (code = 9) to this location      	|
|     1    	| Respondent's home                        	|   1992   	| Walking (code = 7) and Cycling (code = 9) to this location      	|
|     2    	| Work place                               	|   1992   	| Walking (code = 7) and Cycling (code = 9) to this location      	|
|     3    	| Someone else's home                      	|   1992   	| Walking (code = 7) and Cycling (code = 9) to this location      	|
|    4     	| Other place                              	|   1992   	| Walking (code = 7) and Cycling (code = 9) to this location      	|
|     1    	| Respondent's home                        	|   1886   	| Walking (code = 5) to this location                             	|
|     2    	| Work place                               	|   1886   	| Walking (code = 5) to this location                             	|
|     3    	| AT other places                          	|   1886   	| Walking (code = 5) to this location                             	|
|     7    	| other                                    	|   1886   	| Walking (code = 5) to this location                             	|

### What are the bootstrap weights?

Bootstrap weights are only available for 2015? EXPLAIN

```{r}
gss_e_2015 |> 
  slice_head() |>
  select(WGHT_EPI, starts_with("WEPI")) |>
  pivot_longer(cols = -WGHT_EPI, 
               names_to = "Bootstrap_number",
               values_to = "Bootstrap_weight") |>
  ggplot() +
  geom_histogram(aes(x = Bootstrap_weight)) +
  geom_vline(aes(xintercept = WGHT_EPI))
```

WE don't plan to use the bootstrap weights.

## Reading Source Files

In this notebook I experiment with reading the files and then saving them to an R native format for ease of use.


Read main files of GSS survey (Episode Files) from 1986 to 2015:
```{r}
gss_e_2015 <- read.csv(paste0(here(), "/data-inputs/source-files/Time use_2015/gss-e.csv"))
gss_e_2010 <- read.csv(paste0(here(), "/data-inputs/source-files/Time use_2010/gss-12M0018-E-2010-c-24-tus-ef_F1.csv"))
gss_e_2005 <- read.csv(paste0(here(), "/data-inputs/source-files/Time use_2005/gss-12M0019-E-2005-c-19-e_F1.csv"))
gss_e_1998 <- read.csv(paste0(here(), "/data-inputs/source-files/Time use_1998/gss-12M0012-E-1998-c-12e_F1.csv"))
gss_e_1992 <- read.csv(paste0(here(), "/data-inputs/source-files/Time use_1992/gss-12M0007-E-1992-c-7-ep_F1.csv"))
gss_e_1986 <- read.csv(paste0(here(), "/data-inputs/source-files/Time use_1986/gss-12M0002-E-1986-c-2-ep_F1.csv"))
```

We begin by creating a data set for walking using the 2015 data table.

Create a table with origins and destinations for walking in 2015. First, I'll identify the rows in the episode file where the location of the activity was "315", which is code for "walking," as well as one row before and after them to identify the locations before and after the walking that will be considered as the origin and destination for each respondent. 

```{r creating dataset for walking 2015, include=FALSE, cache=FALSE}

# **walking 2015**
# Creating data of origins and destinations of walking trip 2015
inds <- which(gss_e_2015$LOCATION == 315)
rows <- lapply(inds, function(x) (x-1):(x+1))
```

Second, I'll select the required columns (required variables such as PUMFID (Record identification), EPINO (Sequential episode number), WGHT_EPI (Episode weight), DDAY(Diary - Reference day), TOTEPISO (Total number of episodes during the reference day), TUI_01 (Activity code of the episode), STARTIME (Start time of the episode), ENDTIME (End time of the episode), STARTMIN (Start time of the episode in minutes), ENDMIN (End time of the episode in minutes), DURATION (Duration (in minutes) of the episode) and LOCATION (Location of the episode) and then I'll create origins and destinations columns based on the locations before and after the walking for each respondent.

```{r}
walking_2015 <- gss_e_2015 |>
  dplyr::slice(unlist(rows)) |>
  dplyr::select(PUMFID:LOCATION) |> 
  #dplyr::slice_head(n = 15) |>
  group_by(PUMFID) |>
  mutate(origin = lag(LOCATION),
         destination = lead(LOCATION)) |>
  #  group_by(PUMFID) %>% 
  filter(LOCATION == 315) |> 
  ungroup()
```


Similarly, I'll filter origins of interest and then create a column with a description of each origin's code:
```{r}
walking_2015 <- walking_2015 |>
  filter(origin == 300 | 
           origin == 301 | 
           origin == 303 | 
           origin == 302 | 
           origin == 304 |
           origin == 305 | 
           origin == 306 | 
           origin == 307 |
           origin == 308 |
           origin == 309 | 
           origin == 310 |
           origin == 311) |>
  mutate(orig_label =
           case_when(origin == 300 ~ "Home", 
                     origin == 301 ~ "Work or school",
                     origin == 303 ~ "Other's home",
                     origin == 302 ~ "Business",
                     origin == 304 ~ "In the neighbourhood",
                     origin == 305 ~ "Outdoors",
                     origin == 306 ~ "Grocery store, other stores or mall",
                     origin == 308 ~ "Sports centre, field or arena",
                     origin == 307 ~ "Library, museum or theatre",
                     origin == 309 ~ "Restaurant, bar or club",
                     origin == 310 ~ "Place of worship",
                     origin == 311 ~ "Medical, dental or other health clinic"),
         orig_label = factor(orig_label))
```

Then, I'll filter destinations of interest and then create a column with a description of each destination's code
```{r}
# change destination and origins to text column
walking_2015 <- walking_2015 |>  
  # Choose destinations of interest
  filter(destination == 300 | 
           destination == 301 | 
           destination == 303 |
           destination == 304 |
           destination == 302 | 
           destination == 305 |
           destination == 306 | 
           destination == 307 |
           destination == 308 |
           destination == 309 | 
           destination == 310 |
           destination == 311)  |>
  # Label the destinations
  mutate(dest_label =
           case_when(destination == 300 ~ "Home", 
                     destination == 301 ~ "Work or school",
                     destination == 303 ~ "Other's home",
                     destination == 304 ~ "In the neighbourhood",
                     destination == 302 ~ "Business",
                     destination == 305 ~ "Outdoors",
                     destination == 306 ~ "Grocery store, other stores or mall",
                     destination == 307 ~ "Library, museum or theatre",
                     destination == 308 ~ "Sports centre, field or arena",
                     destination == 309 ~ "Restaurant, bar or club",
                     destination == 310 ~ "Place of worship",
                     destination == 311 ~ "Medical, dental or other health clinic"),
         dest_label = factor(dest_label))
```

Here, I'll add two columns of year and mode to walking_2015 dataset.
```{r}
walking_2015 <- walking_2015 |>
  dplyr::select(PUMFID, WGHT_EPI,TUI_01, DDAY, STARTIME:dest_label) |>
  mutate(YEAR = 2015,
         MODE = "Walking")
```


Sanity check!
```{r}
walking_2015 |> 
  group_by(orig_label, 
           dest_label) |>
  count() |>
  pivot_wider(names_from = "dest_label", values_from = "n")
```

The most common destination for a walking trip that starts at home is `Work or school`. Notice, though that there are quite a few trips that start _and_ end at Home. What are those supposed to be?

Check those home-to-home trips:
```{r}
walking_2015 |>
  filter(orig_label == "Home" & dest_label == "Home")
```

See PUMFID == 10041 and check the walking episode:
```{r}
gss_e_2015 |>
  filter(PUMFID == 10041) |>
  slice(unique(c(which(LOCATION == 315) - 1, 
                 which(LOCATION == 315),
                 which(LOCATION == 315) + 1))) |>
  select(PUMFID:LOCATION)
```

The person was at home doing 27 (personal care), then went for walk that lasted 15 minutes, came back home and spent 180 doing 9 (looking for work). So, this was a recreational/leisure trip!

QUESTION: Should we include recreational trips in our analysis? If not, then we need to filter out all home-to-home trips.


Save data for later use:
```{r}
save(walking_2015,
     file = paste0(here(), "/data/walking_2015.Rda"))
```


In the next step, We begin by creating a data set for cycling using the 2015 data table.

Create a table with origins and destinations for cycling in 2015. First, I'll identify the rows in the episode file where the location of the activity was "318", which is code for "cycling," as well as one row before and after them to identify the locations before and after the cycling that will be considered as the origin and destination for each respondent.

```{r creating dataset for cycling 2015, include=FALSE, cache=FALSE}

# cycling _2015
# Creating data of origins and destinations of cycling trip 2015
inds = which(gss_e_2015$LOCATION == 318)
rows <- lapply(inds, function(x) (x-1):(x+1))
```

Second, I'll select the required columns (required variables such as PUMFID (Record identification), EPINO (Sequential episode number), WGHT_EPI (Episode weight), DDAY(Diary - Reference day), TOTEPISO (Total number of episodes during the reference day), TUI_01 (Activity code of the episode), STARTIME (Start time of the episode), ENDTIME (End time of the episode), STARTMIN (Start time of the episode in minutes), ENDMIN (End time of the episode in minutes), DURATION (Duration (in minutes) of the episode) and LOCATION (Location of the episode) and then I'll create origins and destinations columns based on the locations before and after the cycling for each respondent.

```{r}
cycling_2015 <- gss_e_2015[unlist(rows),] %>% 
  dplyr::select(PUMFID:LOCATION) %>% 
  mutate(origin = lag(LOCATION, order_by = PUMFID)) %>% 
  mutate(destination = lead(LOCATION, order_by = PUMFID)) %>% 
  group_by(PUMFID) %>% 
  filter(LOCATION == 318)
```

Similarly, I'll filter origins of interest and then create a column with a description of each origin's code:

```{r}
cycling_2015 <- cycling_2015 |>
  filter(origin == 300 | 
           origin == 301 | 
           origin == 303 | 
           origin == 302 | 
           origin == 304 |
           origin == 305 | 
           origin == 306 | 
           origin == 307 |
           origin == 308 |
           origin == 309 | 
           origin == 310 |
           origin == 311) |>
  mutate(orig_label =
           case_when(origin == 300 ~ "Home", 
                     origin == 301 ~ "Work or school",
                     origin == 303 ~ "Other's home",
                     origin == 302 ~ "Business",
                     origin == 304 ~ "In the neighbourhood",
                     origin == 305 ~ "Outdoors",
                     origin == 306 ~ "Grocery store, other stores or mall",
                     origin == 308 ~ "Sports centre, field or arena",
                     origin == 307 ~ "Library, museum or theatre",
                     origin == 309 ~ "Restaurant, bar or club",
                     origin == 310 ~ "Place of worship",
                     origin == 311 ~ "Medical, dental or other health clinic"),
         orig_label = factor(orig_label))
```

Then, I'll filter destinations of interest and then create a column with a description of each destination's code
```{r}
# change destination and origins to text column
cycling_2015 <- cycling_2015 |>  
  # Choose destinations of interest
  filter(destination == 300 | 
           destination == 301 | 
           destination == 303 |
           destination == 304 |
           destination == 302 | 
           destination == 305 |
           destination == 306 | 
           destination == 307 |
           destination == 308 |
           destination == 309 | 
           destination == 310 |
           destination == 311)  |>
  # Label the destinations
  mutate(dest_label =
           case_when(destination == 300 ~ "Home", 
                     destination == 301 ~ "Work or school",
                     destination == 303 ~ "Other's home",
                     destination == 304 ~ "In the neighbourhood",
                     destination == 302 ~ "Business",
                     destination == 305 ~ "Outdoors",
                     destination == 306 ~ "Grocery store, other stores or mall",
                     destination == 307 ~ "Library, museum or theatre",
                     destination == 308 ~ "Sports centre, field or arena",
                     destination == 309 ~ "Restaurant, bar or club",
                     destination == 310 ~ "Place of worship",
                     destination == 311 ~ "Medical, dental or other health clinic"),
         dest_label = factor(dest_label))
```


Here, I'll add two columns of year and mode to cycling_2015 dataset.

```{r}
cycling_2015 <- cycling_2015 |>
  dplyr::select(PUMFID, WGHT_EPI,TUI_01, DDAY, STARTIME:dest_label) |>
  mutate(YEAR = 2015,
         MODE = "cycling")
```

Sanity check!
```{r}
cycling_2015 |> 
  group_by(orig_label, 
           dest_label) |>
  count() |>
  pivot_wider(names_from = "dest_label", values_from = "n")
```

The most common destination for a cycling trip that starts at home is `Work or school`. Notice, though that there are quite a few trips that start _and_ end at Home. What are those supposed to be?

Check those home-to-home trips:
```{r}
cycling_2015 |>
  filter(orig_label == "Home" & dest_label == "Home")
```
The person was at home doing 27 (personal care), then went for walk that lasted 15 minutes, came back home and spent 180 doing 9 (looking for work). So, this was a recreational/leisure trip!

QUESTION: Should we include recreational trips in our analysis? If not, then we need to filter out all home-to-home trips.





Save data for later use:
```{r}
save(cycling_2015,
    file = paste0(here(), "/data/cycling_2015.Rda"))
```



```{r creating dataset for walking 2005, include=FALSE, cache=FALSE}

# **walking 2005**
# Creating data of origins and destinations of walking trip 2005
inds = which(gss_e_2005$PLACE == 14)
rows <- lapply(inds, function(x) (x-1):(x+1))
walking_2005 <- gss_e_2005 [unlist(rows),] %>% 
  dplyr::select(RECID:PLACE) %>% 
  mutate(origin = lag(PLACE, order_by = RECID)) %>% 
  mutate(destination = lead(PLACE, order_by = RECID)) %>% 
  group_by(RECID) %>% 
  filter(PLACE == 14) %>% 
  ungroup()

# change destination and origins to text column
walking_2005 <- walking_2005 %>%  filter(destination == 1 | destination == 2 | destination == 3 | destination == 8 | destination == 4 | destination == 5 | destination == 6 | destination == 7 | destination == 9 | destination == 10 ) %>% filter(origin == 1 | origin == 2 | origin == 3 | origin == 8 | origin == 4 | origin == 5 | origin == 6 | origin == 7 | origin == 9 | origin == 10) %>% 
  mutate(dest =
           case_when(destination == 1 ~ "home", 
                     destination == 2 ~ "work or school",
                     destination == 3 ~ "other's home",
                     destination == 8 ~ "work or school",
                     destination == 4 ~ "Restaurant, bar or club",
                     destination == 5 ~ "Place of worship",
                     destination == 6 ~ "Grocery store, other stores or mall",
                     destination == 7 ~ "Grocery store, other stores or mall",
                     destination == 9 ~ "outdoors",
                     destination == 10 ~ "Library, museum or theatre")) %>% 
  mutate(orig =
           case_when(origin == 1 ~ "home", 
                     origin == 2 ~ "work or school",
                     origin == 3 ~ "other's home",
                     origin == 8 ~ "work or school",
                     origin == 4 ~ "Restaurant, bar or club",
                     origin == 5 ~ "Place of worship",
                     origin == 6 ~ "Grocery store, other stores or mall",
                     origin == 7 ~ "Grocery store, other stores or mall",
                     origin == 9 ~ "outdoors",
                     origin == 10 ~ "Library, museum or theatre"))

walking_2005 <- walking_2005 %>% dplyr::rename(
  PUMFID = RECID,
  STARTMIN = STARMIN,
  LOCATION = PLACE)

walking_2005 <- walking_2005 %>%  dplyr::select(PUMFID, WGHT_EPI, DDAY, STARTIME:orig) 
walking_2005$YEAR <-  2005
walking_2005$MODE <-  "walking"
```

Save data for later use:
```{r}
save(walking_2005,
     file = paste0(here(), "/data/walking_2005.Rda"))
```


```{r creating dataset for cycling 2005, include=FALSE, cache=FALSE}

# **CYCLING 2005**
# Creating data of origins and destinations of cycling trip 2005
inds = which(gss_e_2005$PLACE == 17)
rows <- lapply(inds, function(x) (x-1):(x+1))
cycling_2005 <- gss_e_2005 [unlist(rows),] %>% 
  dplyr::select(RECID:PLACE) %>% 
  mutate(origin = lag(PLACE, order_by = RECID)) %>% 
  mutate(destination = lead(PLACE, order_by = RECID)) %>% 
  group_by(RECID) %>% 
  filter(PLACE == 17) %>% 
  ungroup()

# change destination and origins to text column
cycling_2005 <- cycling_2005 %>%  filter(destination == 1 | destination == 2 | destination == 3 | destination == 8 ) %>% filter(origin == 1 | origin == 2 | origin == 3 | origin == 8 ) %>% 
  mutate(dest =
           case_when(destination == 1 ~ "home", 
                     destination == 2 ~ "work or school",
                     destination == 3 ~ "other's home",
                     destination == 8 ~ "work or school")) %>% 
  mutate(orig =
           case_when(origin == 1 ~ "home", 
                     origin == 2 ~ "work or school",
                     origin == 3 ~ "other's home",
                     origin == 8 ~ "work or school"))

cycling_2005 <- cycling_2005 %>% dplyr::rename(
  PUMFID = RECID,
  STARTMIN = STARMIN,
  LOCATION = PLACE) %>%  dplyr::select(PUMFID, WGHT_EPI, DDAY, STARTIME:orig) 
cycling_2005$YEAR <-  2005
cycling_2005$MODE <-  "cycling"
```

Save data for later use:
```{r}
save(cycling_2005,
     file = paste0(here(), "/data/cycling_2005.Rda"))
```


```{r creating dataset for walking 1998, include=FALSE, cache=FALSE}

# **walking 1998**
# Creating data of origins and destinations of walking trip 1998
inds = which(gss_e_1998$PLACE == 7)
rows <- lapply(inds, function(x) (x-1):(x+1))
walking_1998 <- gss_e_1998 [unlist(rows),] %>% 
  dplyr::select(RECID:PLACE) %>% 
  mutate(origin = lag(PLACE, order_by = RECID)) %>% 
  mutate(destination = lead(PLACE, order_by = RECID)) %>% 
  group_by(RECID) %>% 
  filter(PLACE == 7) %>% 
  ungroup()

# change destination and origins to text column
walking_1998 <- walking_1998 %>%  filter(destination == 1 | destination == 2 | destination == 3) %>% filter(origin == 1 | origin == 2 | origin == 3) %>% 
  mutate(dest =
           case_when(destination == 1 ~ "home", 
                     destination == 2 ~ "work or school",
                     destination == 3 ~ "other's home")) %>% 
  mutate(orig =
           case_when(origin == 1 ~ "home", 
                     origin == 2 ~ "work or school",
                     origin == 3 ~ "other's home"))

walking_1998 <- walking_1998 %>% dplyr::rename(
  PUMFID = RECID,
  STARTMIN = STARMIN,
  WGHT_EPI = WGHTEPI,
  LOCATION = PLACE) %>%  dplyr::select(PUMFID, WGHT_EPI, DDAY, STARTIME:orig) 
walking_1998$YEAR <-  1998
walking_1998$MODE <-  "walking"
```


Save data for later use:
```{r}
save(walking_1998,
     file = paste0(here(), "/data/walking_1998.Rda"))
```



```{r creating dataset for cycling 1998, include=FALSE, cache=FALSE}

# **Cycling 1998**
# Creating data of origins and destinations of cycling trip 1998
inds = which(gss_e_1998$PLACE == 9)
rows <- lapply(inds, function(x) (x-1):(x+1))
cycling_1998 <- gss_e_1998 [unlist(rows),] %>% 
  dplyr::select(RECID:PLACE) %>% 
  mutate(origin = lag(PLACE, order_by = RECID)) %>% 
  mutate(destination = lead(PLACE, order_by = RECID)) %>% 
  group_by(RECID) %>% 
  filter(PLACE == 9) %>% 
  ungroup()

# change destination and origins to text column
cycling_1998 <- cycling_1998 %>%  filter(destination == 1 | destination == 2 | destination == 3) %>% filter(origin == 1 | origin == 2 | origin == 3) %>% 
  mutate(dest =
           case_when(destination == 1 ~ "home", 
                     destination == 2 ~ "work or school",
                     destination == 3 ~ "other's home")) %>% 
  mutate(orig =
           case_when(origin == 1 ~ "home", 
                     origin == 2 ~ "work or school",
                     origin == 3 ~ "other's home"))

cycling_1998 <- cycling_1998 %>% dplyr::rename(
  PUMFID = RECID,
  STARTMIN = STARMIN,
  WGHT_EPI = WGHTEPI,
  LOCATION = PLACE) %>%  dplyr::select(PUMFID, WGHT_EPI, DDAY, STARTIME:orig) 
cycling_1998$YEAR <-  1998
cycling_1998$MODE <-  "cycling"
```


Save data for later use:
```{r}
save(cycling_1998,
     file = paste0(here(), "/data/cycling_1998.Rda"))
```


```{r creating dataset for walking 1992, include=FALSE, cache=FALSE}
# **walking 1992**
# Creating data of origins and destinations of walking trip 1992
inds = which(gss_e_1992$PLACE == 7)
rows <- lapply(inds, function(x) (x-1):(x+1))
walking_1992 <- gss_e_1992 [unlist(rows),] %>% 
  dplyr::select(SEQNUM, DDAY, NOEPISO, ACTCODE, STARTIME, ENDTIME, DURATION, PLACE, TIMEWGT) %>% 
  mutate(origin = lag(PLACE, order_by = SEQNUM)) %>% 
  mutate(destination = lead(PLACE, order_by = SEQNUM)) %>% 
  group_by(SEQNUM) %>% 
  filter(PLACE == 7) %>% 
  ungroup()

# change NA's Weight to 1 


# change destination and origins to text column
walking_1992 <- walking_1992 %>%  filter(destination == 1 | destination == 2 | destination == 3) %>% filter(origin == 1 | origin == 2 | origin == 3) %>% 
  mutate(dest =
           case_when(destination == 1 ~ "home", 
                     destination == 2 ~ "work or school",
                     destination == 3 ~ "other's home")) %>% 
  mutate(orig =
           case_when(origin == 1 ~ "home", 
                     origin == 2 ~ "work or school",
                     origin == 3 ~ "other's home"))

walking_1992 <- walking_1992 %>% dplyr::rename(
  PUMFID = SEQNUM,
  LOCATION = PLACE, WGHT_EPI = TIMEWGT) %>%  dplyr::select(PUMFID, DDAY, STARTIME:orig) 
walking_1992$YEAR <-  1992
walking_1992$MODE <-  "walking"
walking_1992$STARTMIN <-  1
walking_1992$ENDMIN <-  1
```

Save data for later use:
```{r}
save(walking_1992,
     file = paste0(here(), "/data/walking_1992.Rda"))
```


```{r creating dataset for cycling 1992, include=FALSE, cache=FALSE}

# **Cycling 1992**
# Creating data of origins and destinations of cycling trip 1992
inds = which(gss_e_1992$PLACE == 9)
rows <- lapply(inds, function(x) (x-1):(x+1))
cycling_1992 <- gss_e_1992 [unlist(rows),] %>% 
  dplyr::select(SEQNUM, DDAY, NOEPISO, ACTCODE, STARTIME, ENDTIME, DURATION, PLACE, TIMEWGT) %>% 
  mutate(origin = lag(PLACE, order_by = SEQNUM)) %>% 
  mutate(destination = lead(PLACE, order_by = SEQNUM)) %>% 
  group_by(SEQNUM) %>% 
  filter(PLACE == 9) %>% 
  ungroup()

# change destination and origins to text column
cycling_1992 <- cycling_1992 %>%  filter(destination == 1 | destination == 2 | destination == 3) %>% filter(origin == 1 | origin == 2 | origin == 3) %>% 
  mutate(dest =
           case_when(destination == 1 ~ "home", 
                     destination == 2 ~ "work or school",
                     destination == 3 ~ "other's home")) %>% 
  mutate(orig =
           case_when(origin == 1 ~ "home", 
                     origin == 2 ~ "work or school",
                     origin == 3 ~ "other's home"))

cycling_1992 <- cycling_1992 %>% dplyr::rename(
  PUMFID = SEQNUM,
  LOCATION = PLACE,WGHT_EPI = TIMEWGT) %>%  dplyr::select(PUMFID, DDAY, STARTIME:orig) 
cycling_1992$YEAR <-  1992
cycling_1992$MODE <-  "cycling"
cycling_1992$STARTMIN <-  1
cycling_1992$ENDMIN <-  1

```

Save data for later use:
```{r}
save(cycling_1992,
     file = paste0(here(), "/data/cycling_1992.Rda"))
```

```{r creating dataset for walking 1986, include=FALSE, cache=FALSE}

# **walking 1986**
# Creating data of origins and destinations of walking trip 1986
inds = which(gss_e_1986$PLACE == 7)
rows <- lapply(inds, function(x) (x-1):(x+1))
walking_1986 <- gss_e_1986 [unlist(rows),] %>% 
  dplyr::select(SEQNUM, DDAY, NO_EPISO, ACT_CODE, STRTTIME, ENDTIME, DURATION, PLACE, FWGT_MS) %>% 
  mutate(origin = lag(PLACE, order_by = SEQNUM)) %>% 
  mutate(destination = lead(PLACE, order_by = SEQNUM)) %>% 
  group_by(SEQNUM) %>% 
  filter(PLACE == 7) %>% 
  ungroup()
summary(walking_1986$PLACE)

# change destination and origins to text column
walking_1986 <- walking_1986 %>%  filter(destination == 1 | destination == 2 | destination == 3) %>% filter(origin == 1 | origin == 2 | origin == 3) %>% 
  mutate(dest =
           case_when(destination == 1 ~ "home", 
                     destination == 2 ~ "work or school",
                     destination == 3 ~ "other's home")) %>% 
  mutate(orig =
           case_when(origin == 1 ~ "home", 
                     origin == 2 ~ "work or school",
                     origin == 3 ~ "other's home"))

walking_1986 <- walking_1986 %>% dplyr::rename(
  PUMFID = SEQNUM,
  LOCATION = PLACE, WGHT_EPI = FWGT_MS, STARTIME = STRTTIME) %>%  dplyr::select(PUMFID, DDAY, STARTIME:orig) 
walking_1986$YEAR <-  1986
walking_1986$MODE <-  "walking"
walking_1986$STARTMIN <-  1
walking_1986$ENDMIN <-  1
```

Save data for later use:
```{r}
save(walking_1986,
     file = paste0(here(), "/data/walking_19869.Rda"))
```




